Raft Distributed Consensus Algorithm over WebRTC (Web Real-Time Communication)

### 2015-08-23 research_class/paper.tex
Expand outcome section. Comment out TODOs

The Raft algorithm breaks the problem of consensus into three core
subproblems: leader election, log replication, and safety. A full Raft
solution must also address: cluster membership changes, log
compaction, and client interaction. The Raft algorithm implemented and
discussed in this paper is based on "Consensus: Bridging Theory and
Practice" by Diego Ongaro (PhD Dissertation) TODO/CITE.

@@

Raft.js is an implementation of the Raft algorithm in JavaScript that
was creator by the author of this paper TODO/CITE. Raft.js is designed
to run in either a browser environment or within node.js (server-side
JavaScript). Raft.js implements the Raft algorithm as described in
"Consensus", Ongaro (dissertation TODO/CITE).

@@

The WebRTC browser APIs enable web applications written in JavaScript
to establish browser-to-browser network connectivity in order to
stream real-time video, audio and data communication. These APIs are
currently in the process of being standardized by the W3C organization
(TODO/CITE: http://www.w3.org/TR/webrtc/)

@@

The PeerJS project provides a simple WebRTC signaling server called
PeerServer. This server uses HTTP and WebSockets transport protocols
to perform WebRTC signaling on behalf of browser clients. The
PeerServer server is also extensible so that more advanced services
can be built with it. The PeerJS project includes an extension to
PeerServer called ExpressPeerServer that combines a PeerJS signaling
server with the Node.js based Express web application framework
(TODO/CITE).

@@

The web application also begins a periodic async polling function
(`addRemoveServersAsync`) when the page loads. When the node is a Raft
leader, the function compares the current active list of PeerJS clients to
the Raft cluster node list. If there are any differences between the
two lists, then any new clients are added to the Raft cluster and
missing clients are removed from the Raft cluster one at a time using
the `addServer` and `removeServer` RPC calls. Servers additions are
prioritized before removals (as recommended in Ongaro 4.4 TODO/CITE)
in order to maximize availability.

@@

\subsubsection{Client Request Implementation}

TODO: client request functionality

@@

The chat application extends the basic Raft.js over WebRTC client. The
chat application provides a large text area containing the chat
history and a text entry field below it. When the user clicks on the
\emph{Send} button (or pressed enter), the text line is added to
a pending send queue and the text entry field is blanked. A function
to flush sends runs periodically (currently every 100ms) and does the
following:

@@

TODO: screenshot.

TODO: atomic seqAppend operation

TODO: longer delay for propagation in chat application because each
addition to chat history take several messages to fully propagate.

@@

In addition, the full chat application was also tested manually. TODO:
more text.

@@

@@

TODO: image of chat app with one node

TODO: image of chat app five nodes

@@

@@

TODO: Setup for the automated tests. Description of SlimerJS. Creation of SlimerJS test scripts using JavaScript.

@@

TODO: Outcome of the automated tests.

The following figure shows the amount of time that the cluster takes to
elect a leader and then reconfigure the cluster to the the final size
(adding one node a time).

The following figure shows the amount of time that it takes for
a stable cluster to recover (elect a new leader and remove the failed
server from the cluster) when the current leader fails.

The following figure shows the amount of time that it takes for
a stable cluster to recover (elect a new leader and remove the failed
servers from the cluster) when one or more nodes fails. The first line
shows the recovery time when just the leader fails. The second line
show the recovery time when one less than half of the nodes fail.

\begin{figure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et100={mark=x,draw=blue},
                    et100_fail={mark=x,draw=red},
                    et100_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{100ms election timer}
        \label{fig:A}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et500={mark=x,draw=blue},
                    et500_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{500ms election timer}
        \label{fig:B}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et1000={mark=x,draw=blue},
                    et1000_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{1000ms election timer}
        \label{fig:C}
    \end{subfigure}
    \caption{Initial Cluster Activation}
    \label{fig:cluster_up}

@@

    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5, xmax=23,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{100ms election timer}
        \label{fig:G}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{500ms election timer}
        \label{fig:H}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{1000ms election timer}
        \label{fig:I}
    \end{subfigure}
    \caption{Recovery from Half Cluster Failure}
    \label{fig:G}
\end{figure}

The following figure shows the amount of time that a stable cluster takes to
propagate, commit and apply a single state machine command to the state
machines on every cluster node.

>>>>

The Raft algorithm breaks the problem of consensus into three core
subproblems: leader election, log replication, and safety. A full Raft
solution must also address: cluster membership changes, log
compaction, and client interaction. The Raft algorithm implemented and
discussed in this paper is based on "Consensus: Bridging Theory and
Practice" by Diego Ongaro (PhD Dissertation). % TODO/CITE

@@

Raft.js is an implementation of the Raft algorithm in JavaScript that
was creator by the author of this paper. %TODO/CITE
Raft.js is designed to run in either a browser environment or within
node.js (server-side JavaScript). Raft.js implements the Raft
algorithm as described in "Consensus", Ongaro (dissertation).
% TODO/CITE

@@

The WebRTC browser APIs enable web applications written in JavaScript
to establish browser-to-browser network connectivity in order to
stream real-time video, audio and data communication. These APIs are
currently in the process of being standardized by the W3C organization
(http://www.w3.org/TR/webrtc/) % TODO/CITE

@@

The PeerJS project provides a simple WebRTC signaling server called
PeerServer. This server uses HTTP and WebSockets transport protocols
to perform WebRTC signaling on behalf of browser clients. The
PeerServer server is also extensible so that more advanced services
can be built with it. The PeerJS project includes an extension to
PeerServer called ExpressPeerServer that combines a PeerJS signaling
server with the Node.js based Express web application framework.
% (TODO/CITE)

@@

The web application also begins a periodic async polling function
(`addRemoveServersAsync`) when the page loads. When the node is a Raft
leader, the function compares the current active list of PeerJS clients to
the Raft cluster node list. If there are any differences between the
two lists, then any new clients are added to the Raft cluster and
missing clients are removed from the Raft cluster one at a time using
the `addServer` and `removeServer` RPC calls. Servers additions are
prioritized before removals (as recommended in Ongaro 4.4) % TODO/CITE
in order to maximize availability.

@@

\subsubsection{Client Request Redirect Implementation}

In the Raft over WebRTC implemenation each local browser is a client
of the Raft cluster. However in Raft, the leader node is the only node
that adds entries to the log so clients on follower nodes must have
a way to communicate with the leader. This is implemented as pair of
functions named clientRequest and clientRequestResponse. clientRequest
is called by the client with the arguments of a client request RPC and
a callback function to be called once the request is completed. The
clientRequest/clientRequestResponse functions work as follows:

\begin{enumerate}
\item the clientRequestResponse is registered with the local Raft.js
    implementation as the callback function anytime
    clientRequestResponse RPCs are received.
\item the clientRequest function keeps track of the current leader ID.
    If the current leader ID is not set or this node is the leader
    then the client's request is sent directly to the local Raft node.
    If the current leader ID is set then the request is sent to that
    node as a normal RPC request over the WebRTC channel.
\item when the clientRequestResponse function is called and the status
    is "NOT\_LEADER" then the current leader ID being tracked by the
    clientRequest function is set to the leader ID returned in the
    response. Then the original request is re-issued over the normal
    WebRTC channel to the leader.
\item when the clientRequestResponse function is called with
    a successful status, the original callback argument of the
    clientRequest function is called with the results from the RPC
    response.
\end{enumerate}

@@

The chat application extends the basic Raft.js over WebRTC client. The
chat application provides a large text area containing the chat
history and a text entry field below it. When the user clicks on the
\emph{Send} button (or pressed enter), the current node ID is
prepended to the beginning of the text line and it is added to
a pending send queue and the text entry field is blanked. A function
to flush sends runs periodically (currently every 100ms) and does the
following:

@@

The chat application also extends the default set of state machine
operations to add a seqAppend operation. This is an atomic operation
that is implemented as follows:

\begin{enumerate}
\item the seqAppend operation operates on a specially structured
    key/value pair in the state machine map. The value is
    a associative map that contains a counter and list structure that
    contains the actual data. seqAppend operations are sent with both
    a count and then new value to append.
\item When a seqAppend operation is applied to the state machine and
    the count value in the command matches the counter in the state
    machines value, then the new value in the request is added to the
    list and the counter is incremented. If the count value does not
    match then counter then the the operation is not applied. In
    either case, a list is returned that contains two values: whether
    the seqAppend operation succeeded and what the current value of
    the counter.
\item the first time the seqAppend operation is used on a non-existent
    key, the structure described above is created with the counter set
    to 0 and the list structure is created with the single value
    passed in the seqAppend operation.
\item clients that call the seqAppend operation should check the
    status value in the result list. If the status is false, then the
    client should retry the operation using the counter value as the
    count in the operation.
\end{enumerate}

@@

In addition, the full chat application was also tested manually by
loading the initial chat application landing page to create a Raft
over WebRTC cluster with a single leader. Then one or more lines of
text were entered into the text line each followed by clicking the
Send button.  Here is a screenshot of the browser after the first
node/tab of the chat application is created and a line of text was
sent:

@@

Then four more Raft nodes were then added to the cluster by clicking
on the embedded "create more clients" link in the application.
Finally, on a non-leader node more text lines were entered and sent by
clicking the Send button. Here is a screenshot of the browser once
four new nodes/tabs are added to the cluster (5 total) and another
text line was sent:

@@

@@

In the context of the full chat application the slowdown of background
pages is even more pronounced because each addition to the chat
history can take several messages to fully propagate the new date.

@@

Automated testing of the Raft over WebRTC system was performed by
leveraging the SlimerJS scriptable browser environment. SlimerJS
enables web applications to be scripted by loading pages in a special
Mozilla rendering environment called XULRunner. The XULRunner instance
is directed to a X virtual frame buffer (xvfb) which provides
effectively headless testing.

To simplify the execution of the tests and to make the testing
infrastructure more portable, a docker container environment was used
to encapsulate and execute the SlimerJS environment. The container is
built using \emph{test/Dockerfile} which is in turn derived from
the "fentas/slimerjs" docker image
(https://github.com/fentas/docker/blob/master/slimerjs/Dockerfile).
This custom Dockerfile was created because the upstream version of the
Dockerfile used an older version of XULRunner that did not support
more than 7 WebRTC clients communicating together simultaneously.
% (TODO/CITE)

Writing tests using SlimerJS involves creating JavaScript programs
that are run within the SlimerJS environment. These scripts can create
browser instances (basically headless browser tabs), query the status
of those pages and execute JavaScript within the context of the pages
themselves.

Three different tests were created to test the Raft over WebRTC
functionality and the chat application. Each of the tests takes
a parameter indicating the number of nodes/pages to launch and test
again. The tests scripts files are:

\begin{enumerate}
    \item \emph{test/common.js}: a module of common functions used by
        the other tests scripts.
    \item \emph{test/test\_up.js}: launch and start an initial leader
        node/page, then start the rest of the nodes (if any) so that
        the leader will configure them into the cluster. Then test
        then waits for every node to be recorded (applied) into the log
        of every other node of the cluster. The test reports the
        number of milliseconds for the cluster to reach this state.
    \item \emph{test/test\_kill\_nodes.js}: similar to
        \emph{test/test\_up.js} but once all the nodes are configured
        into the cluster, this script then kills the leader plus
        a configurable number of nodes. By default the script kills
        just under half the nodes of the cluster. The test then waits
        for a new leader to be elected and remove the dead nodes from
        the cluster. The test reports both the number of milliseconds
        for the cluster to come up and the recover from the failure.
    \item \emph{test/test\_chat\_propagate.js}: similar to
        \emph{test/test\_up.js} but once the cluster is up, the test
        connects to one of the non-leader nodes and and sends a chat
        text line. The test then waits for the chat message to
        propogate to every other node's state machine. The test
        reports both the number of milliseconds for the cluster to
        come up and for the messages to propagate to every node.
\end{enumerate}

The tests were all run on a system with the following configuration:

\begin{enumerate}
    \item \emph{CPUs}: 8 Intel Core i7 processors running at 3.20GHz
    \item \emph{Memory}: 12 GB
    \item \emph{OS}: Ubuntu 14.04.1 LTS
    \item \emph{Docker}: version 1.3.1
    \item \emph{Docker image}:
    \begin{enumerate}
        \item \emph{OS}: Ubuntu 12.04.5 LTS
        \item \emph{SlimerJS}: vesion 0.9.6
        \item \emph{XULRunner}: 38.0.5
    \end{enumerate}
\end{enumerate}

Three tests were run at 10 different cluster sizes (odd cluster sizes
from 3 to 21 nodes): \emph{test/test\_up.js} and
\emph{test/test\_kill\_nodes.js} in kill one (leader) node and kill
half nodes. Each of these tests was run with three different Raft election
timeout values: 100ms, 500ms, 1000ms. The
\emph{test/test\_chat\_propagate.js} test was run at
11 different cluster sizes (odd cluster sizes from 1 to 21 nodes).
Each test mode and cluster size was run 7 times for a total of 707
test runs (3*10*7*3 + 11*7)

@@

% TODO: programmatic figure numbers
Figure 7 shows the results of running the
\emph{test/test\_chat\_propagate.js} test to show the amount of time
that a stable cluster takes to propagate, commit and apply a single
state machine command to the state machines on every cluster node.

Figure 7 is a scatterplot with each test run as a blue "x" and a black
line representing the average at each cluster size.

Figures 8, 9 and 10 show the results of running the
\emph{test/test\_up.js} and \emph{test/test\_kill\_nodes.js} in both
modes (kill only leader and kill just under half the nodes). Three
different scatterplots are shown for each test mode to test different
values of the Raft election timeout value: 100ms, 500ms, and 1000ms.
The election timeout value makes very little difference to the average
time curve line. The reason for this is that the first leader tends to
stay the leader as new nodes are added so for the most part the
election timer does not have an impact.

Some specific notes about the figures:

% TODO: link figure numbers

\begin{enumerate}
    \item \emph{Figure 8 (a)}: The results show that once the cluster
        size reaches 17, the cluster sometimes does not activate
        before the test times out (the red "x" marks). Once the
        cluster reaches 21 nodes, all 7 test runs time out. The reason
        for this is because at 17 nodes, the leader is not always able
        to send out appendEntries RPCs to all the other nodes before
        their election timeout fires. More analysis is needed, but one
        possible explanation is that the single SlimerJS XULRunner
        process that is handling all the cluster nodes is not able to
        give enough CPU time to the leader node to reliable send
        appendEntries RPCs to all the other nodes within the election
        timeout value. Increasing the election timeout to 500ms or
        greater eliminates the failures without significantly changing
        the average cluster activation time curve.
    \item \emph{Figure 9 (a)}: The results show that when the cluster
        size reaches 15, then average amount of time for cluster
        recovery (after the leader is killed) increases dramatically
        (not all data points are shown on the plot for size 15). At
        cluster size 17 the cluster is no longer able to recover
        within the test timeout period. Increasing the election
        timeout to 500ms or greater eliminates the failures at the
        cost of slightly increased recovery time for smaller cluster
        sizes.
    \item \emph{Figure 10 (a)}: During this test cluster was unable to
        activate at cluster size 21 when the election timeout value is
        set to 100ms (see Figure 8 (a)). This means there was no
        opportunity for cluster recovery at cluster size 21 so there
        is no data for that column.

\end{enumerate}

Each figure is a scatterplot with each test run as a blue "x" and
a black line representing the average at each cluster size.

%\end{multicols}

\begin{figure*}[h]
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et100={mark=x,draw=blue},
                    et100_fail={mark=x,draw=red},
                    et100_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{100ms election timer}
        \label{fig:A}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et500={mark=x,draw=blue},
                    et500_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{500ms election timer}
        \label{fig:B}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et1000={mark=x,draw=blue},
                    et1000_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{1000ms election timer}
        \label{fig:C}
    \end{subfigure}
    \caption{Initial Cluster Activation}
    \label{fig:cluster_up}

@@

    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5, xmax=23,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{100ms election timer}
        \label{fig:G}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{500ms election timer}
        \label{fig:H}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{1000ms election timer}
        \label{fig:I}
    \end{subfigure}
    \caption{Recovery from Half Cluster Failure}
    \label{fig:G}
\end{figure*}

%\begin{multicols}{2}
